|

### 2022-06-14 draft-1.Rmd
Draft Intro References

Proximity to greenspace is a potential determinant on an individual’s mental health [@nutsford2013; @ekkel2017; @collins2020; @sakar2018; @jiang2021]. As a low-cost mental health intervention, greenspace provides restorative experiences, encourages social cohesion, reduces pollution, and offers opportunities for physical activity [@lachowycz2013; @wang2021]. To date, most greenspace research has focused on the greenspace mental health relationship in urban areas, with less attention focused on rural communities [@collins2020]. Within studies that have considered the greenspace mental health relationship in rural areas, evidence is mixed, with some studies finding greenspace benefits mental health similarly in rural areas as compared to urban areas [@verheji2008; @alcock2015]. Other research suggests that the rural greenspace mental health relationship differs from the urban relationship [@jiang2021], with private greenspaces — including yards, agricultural fields, and general “countryside greenness” — playing an important role in the rural greenspace mental health relationship [@ekkel2017]. As such, further investigation into the rural greenspace mental health relationship is needed, specifically with regards to public greenspace quality, in addition to greenspace quantity — i.e., available area within a greenspace.

Most greenspace studies consider greenspace quantity as the primary metric, with less attention given to greenspace quality (Collins et al., 2020). Past research has found that an increased quantity of greenspace tends to correspond with better mental health outcomes (van Dillen et al., 2012; Wood et al., 2017; Wang et al., 2021). However, more research is needed to determine whether greenspace quality — attributes including nearby amenities, “greenness” metrics like NDVI, or the presence of public hiking or biking trails — are significant predictors of mental health outcomes.

Of the studies that have considered quality, the most common technique is the use of surveys, allowing the researcher to assess self-reported high-quality greenspace and mental health outcomes (Wang et al. 2021). Other studies have qualified greenspace through consideration of the physical characteristics of the greenspace itself, such as land cover, conservation easements, land designations, species diversity and health and presence of birds (Wheeler et al. 2015), as well as water features and heavy tree cover which provides numerous options for shade in hot months (Mears et al. 2019). The findings from these studies suggest that the quality of greenspace may be a crucial factor in the greenspace mental health relationship. Thus, consideration of greenspace quality is critical in better understanding the greenspace mental health relationship.

Western North Carolina is a predominantly rural, mountainous region. Mental health care access is limited (MAHEC, 2022), as is the case in many rural regions in the United States, despite similar or higher mental health case counts occurring in rural versus urban areas (McCall-Hosenfeld 2014). As such, investigation into low-cost mental health care interventions is important. This study investigates the greenspace mental health relationship in Western North Carolina to understand better the greenspace mental health relationship in rural, mountainous regions. To address research gaps, this study will consider greenspace quality in a region comprising mostly rural communities. This research adds to the growing body of greenspace mental health literature and adds new knowledge by assessing how greenspace quality influences mental health in rural regions.

>>>>

Proximity to greenspace is a potential determinant on an individual’s mental health [@nutsford2013; @ekkel2017; @collins2020; @sarkar2018; @jiang2021]. As a low-cost mental health intervention, greenspace provides restorative experiences, encourages social cohesion, reduces pollution, and offers opportunities for physical activity [@lachowycz2013; @wang2021]. To date, most greenspace research has focused on the greenspace mental health relationship in urban areas, with less attention focused on rural communities [@collins2020]. Within studies that have considered the greenspace mental health relationship in rural areas, evidence is mixed, with some studies finding greenspace benefits mental health similarly in rural areas as compared to urban areas [@verheij2008; @alcock2015]. Other research suggests that the rural greenspace mental health relationship differs from the urban relationship [@jiang2021], with private greenspaces — including yards, agricultural fields, and general “countryside greenness” — playing an important role in the rural greenspace mental health relationship [@ekkel2017]. As such, further investigation into the rural greenspace mental health relationship is needed, specifically with regards to public greenspace quality, in addition to greenspace quantity — i.e., available area within a greenspace.

Most greenspace studies consider greenspace quantity as the primary metric, with less attention given to greenspace quality [@collins2020]. Past research has found that an increased quantity of greenspace tends to correspond with better mental health outcomes [@vandillen2012; @wood2017; @wang2021]. However, more research is needed to determine whether greenspace quality — attributes including nearby amenities, “greenness” metrics like NDVI, or the presence of public hiking or biking trails — are significant predictors of mental health outcomes.

Of the studies that have considered quality, the most common technique is the use of surveys, allowing the researcher to assess self-reported high-quality greenspace and mental health outcomes [@wang2021]. Other studies have qualified greenspace through consideration of the physical characteristics of the greenspace itself, such as land cover, conservation easements, land designations, species diversity and health and presence of birds [@wheeler2015], as well as water features and heavy tree cover which provides numerous options for shade in hot months [@mears2019]. The findings from these studies suggest that the quality of greenspace may be a crucial factor in the greenspace mental health relationship. Thus, consideration of greenspace quality is critical in better understanding the greenspace mental health relationship.

Western North Carolina is a predominantly rural, mountainous region. Mental health care access is limited [@mahec], as is the case in many rural regions in the United States, despite similar or higher mental health case counts occurring in rural versus urban areas [@mccall-hosenfeld2014]. As such, investigation into low-cost mental health care interventions is important. This study investigates the greenspace mental health relationship in Western North Carolina to understand better the greenspace mental health relationship in rural, mountainous regions. To address research gaps, this study will consider greenspace quality in a region comprising mostly rural communities. This research adds to the growing body of greenspace mental health literature and adds new knowledge by assessing how greenspace quality influences mental health in rural regions.
